\pkg{cv}: An \proglang{R} Package for Cross-Validating Regression Models

### 2024-07-11 scratch/notes_john/JSS-article/JSS-article-4.Rmd
make Figs. 6 & 7 larger in JSS-article-4

The \code{Auto} data appeared in a preliminary example in Section \ref{preliminary-example-polynomial-regression}, where we employed CV to inform the selection of the degree of a polynomial regression of \code{mpg} on \code{horsepower}. Here, we consider more generally the problem of predicting \code{mpg} from the other variables in the \code{Auto} data. We begin with a bit of data management, and then examine the pairwise relationships among the numeric variables in the data set (Figure \ref{fig:Auto-explore} (a), produced by the \code{scatterplotMatrix()} function in the \pkg{car} package):
```{r Auto-explore, fig.show="hold", echo=FALSE}
#| out.width = "45%",
#| fig.height = 5,
#| fig.cap = "(a) Scatterplot matrix for the numeric variables in the \\code{Auto} data. In each panel, the solid line shows the linear least-squares fit and the broken line is for a nonparametric regression. (b) Component + residual plots for the working model fit to the \\code{Auto} data. The broken blue lines show linear least-squares fits, while the solid magenta lines are nonparametric-regression smooths.",
#| fig.subcap=rep("", 2)
Auto$cylinders <- factor(Auto$cylinders,
                         labels = c("3-4", "3-4", "5-6", "5-6", "8"))
Auto$year <- as.factor(Auto$year)
Auto$origin <- factor(Auto$origin,
                      labels = c("America", "Europe", "Japan"))
rownames(Auto) <- make.names(Auto$name, unique = TRUE)
Auto$name <- NULL

@@

We start with a "working model" that specifies linear partial relationships of the response to the numeric predictors:
```{r Auto-working-model, eval=FALSE}
m.auto <- lm(mpg ~ ., data = Auto)
crPlots(m.auto, pch = ".")
```
The component + residual plots in Figure \ref{fig:Auto-explore} (b), produced by the \code{cvPlots()} function in the \pkg{car} package, clearly reveal the inadequacy of the model.

@@

The transformed numeric variables are much better-behaved (cf., Figure \ref{fig:Auto-explore} (a), on page \pageref{fig:Auto-explore}, and Figure \ref{fig:Auto-transformed} (a)):
```{r Auto-transformed, fig.show="hold", echo=FALSE}
#| out.width = "45%",
#| fig.height = 5,
#| fig.cap = "(a) Scatterplot matrix for the transformed numeric variables in the \\code{Auto} data. (b) Component + residual plots for the regression model fit to the transformed \\code{Auto} data.",
#| fig.subcap=rep("", 2)
scatterplotMatrix(~ log(mpg) + displacement + horsepower + weight 
                  + acceleration, 
                  smooth=list(spread = FALSE), data = A, pch = ".")
crPlots(m, pch = ".")
```

>>>>

The \code{Auto} data appeared in a preliminary example in Section \ref{preliminary-example-polynomial-regression}, where we employed CV to inform the selection of the degree of a polynomial regression of \code{mpg} on \code{horsepower}. Here, we consider more generally the problem of predicting \code{mpg} from the other variables in the \code{Auto} data. We begin with a bit of data management, and then examine the pairwise relationships among the numeric variables in the data set (Figure \ref{fig:Auto-explore} (a), produced by the \code{scatterplotMatrix()} function in the \pkg{car} package); in each panel, the solid line shows the linear least-squares fit and the broken line is for a nonparametric regression.:
```{r Auto-explore, fig.show="hold", echo=FALSE}
#| out.width = "65%",
#| fig.align = 'center',
#| fig.height = 5,
#| fig.ncol = 1,
#| fig.cap = "(a) Scatterplot matrix for the numeric variables in the \\code{Auto} data.  (b) Component + residual plots for the working model fit to the \\code{Auto} data.",
#| fig.subcap=rep("", 2)
Auto$cylinders <- factor(Auto$cylinders,
                         labels = c("3-4", "3-4", "5-6", "5-6", "8"))
Auto$year <- as.factor(Auto$year)
Auto$origin <- factor(Auto$origin,
                      labels = c("America", "Europe", "Japan"))
rownames(Auto) <- make.names(Auto$name, unique = TRUE)
Auto$name <- NULL

@@

We start with a "working model" that specifies linear partial relationships of the response to the numeric predictors:
```{r Auto-working-model, eval=FALSE}
m.auto <- lm(mpg ~ ., data = Auto)
crPlots(m.auto, pch = ".")
```
The component + residual (C+R) plots in Figure \ref{fig:Auto-explore} (b), produced by the \code{cvPlots()} function in the \pkg{car} package, clearly reveal the inadequacy of the model. The broken blue lines in the C+R plots show linear least-squares fits, while the solid magenta lines are nonparametric-regression smooths.

@@

The transformed numeric variables are much better-behaved (cf., Figure \ref{fig:Auto-explore} (a), on page \pageref{fig:Auto-explore}, and Figure \ref{fig:Auto-transformed} (a)):
```{r Auto-transformed, fig.show="hold", echo=FALSE}
#| out.width = "65%",
#| fig.align = 'center',
#| fig.height = 5,
#| fig.ncol = 1,
#| fig.cap = "(a) Scatterplot matrix for the transformed numeric variables in the \\code{Auto} data. (b) Component + residual plots for the regression model fit to the transformed \\code{Auto} data.",
#| fig.subcap=rep("", 2)
scatterplotMatrix(~ log(mpg) + displacement + horsepower + weight 
                  + acceleration, 
                  smooth=list(spread = FALSE), data = A, pch = ".")
crPlots(m, pch = ".")
```
